greta.Rmd

### 2019-10-09 notebook/appendix.Rmd
apply h first much cleaner

damage <- 0.05
control <- 1
discount <- 0.98

n_s <- 110
states <- seq(0,110, length=n_s)
actions <- states

p <- list(r = .8, K = 153, q = 2, b = 20, sigma = .02, a=2.3, x0 = 20)
eps <- states[2] / 10

may <- function(a){  
  function(x, h = 0){ # May
    pmax(
      x + x * p$r * (1 - x / p$K)  - a * x ^ p$q / (x ^ p$q + p$b ^ p$q) + eps - h,  
      0)
  }
}

reward_fn <- function(x,h) - damage * (x / 100) - control * h / 100

@@

true_a <- 27.4      ### 21.5
believe_a <- 29 # 28.5   ### 22.5

@@

transition_matrix <- function(states, actions, f, sigma){
    n_s <- length(states)
    n_a <- length(actions)
    transition <- array(0, dim = c(n_s, n_s, n_a))
    for(i in 1:n_a){
      for (k in 1:n_s) {
          nextpop <- f(states[k], actions[i])
          if (nextpop < 0) {
          #if(which.min(abs(states - nextpop)) > 1) {
             x  <- c(1, rep(0, n_s - 1))
          } else {
            x <- truncnorm::dtruncnorm(states, 0, max(states), nextpop, sigma * nextpop)
            if(sum(x) <= 0){ 
              x  <- c(1, rep(0, n_s - 1))
            } else {
              x <- x / sum(x)
            }
          }
          transition[k, , i] <- x
      }
  }
  if(any(is.na(transition))) stop("error creating transition matrix")
    transition
}
```

@@

@@

prob_dynamics(m_true[,,1], X, Tmax) %>% 
  ggplot(aes(state,probability))  + geom_point() # + geom_bar(stat="identity")
```

@@

```{r}
x0 <- which.min(abs(states - 23))
```

```{r}
no_policy <- numeric(length(states)) + 1

#set.seed(12345678)
set.seed(12345)

df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = no_policy, x0 = x0, Tmax = Tmax)

```{r}
tibble(state = states,
       belief = actions[belief_p$policy],
       ideal =  actions[true_p$policy]) %>%
  pivot_longer(c(belief, ideal), 
               names_to="model",
               values_to = "action") %>%
  ggplot(aes(state,action, color = model)) + geom_point() 
```

```{r}
set.seed(12345)
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = x0, Tmax = 1000)

```{r}
set.seed(12345)
df <- mdp_planning(m_belief, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = 26, Tmax = 1000)

df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = true_p$policy, x0 = x0, Tmax = 1000)

```{r}
if(!file.exists("learning_sim.tsv.gz")){ # Cache the slow step

  learning_sim <- mdp_learning(transition, reward, discount, 
                      x0 = x0, 
                      Tmax = Tmax, 
                      true_transition = transition[[true_i]],
                      model_prior = prior,
                      type = "value iteration", 
                      epsilon = 1e-2)
  readr::write_tsv(learning_sim, "learning_sim.tsv.gz")

} else {
  learning_sim <- readr::read_tsv("learning_sim.tsv.gz")
}
```

>>>>

damage <- 0.05
control <- 1
discount <- 0.999

n_s <- 121
states <- seq(0, 120, length = n_s)
actions <- seq(0, 120, length = n_s)

p <- list(r = .8, K = 153, q = 2, b = 20, sigma = .02, a=2.3, x0 = 20)

may <- function(a){  
  function(x, h = 0){ # May
    y <- x - h
    pmax(
      ## controlling h is controlling the bifurcation point directly...
      y + y * p$r * (1 - y / p$K)  - a * y ^ p$q / (y ^ p$q + p$b ^ p$q),  
      0)
  }
}

reward_fn <- function(x,h) - damage * (x / 100) ^ 2 - control * (h / 100)

@@

true_a <- 27.2      ### 21.5
believe_a <- 29 # 28.5   ### 22.5

@@

transition_matrix <- function(states, actions, f, sigma){
    n_s <- length(states)
    n_a <- length(actions)
    transition <- array(0, dim = c(n_s, n_s, n_a))
    for(i in 1:n_a){
      for (k in 1:n_s) {
          nextpop <- f(states[k], actions[i])
          if (nextpop <= 0) {
          #if(which.min(abs(states - nextpop)) > 1) {
             x  <- c(1, rep(0, n_s - 1))
          } else {
            x <- truncnorm::dtruncnorm(states, 0, max(states), nextpop, sigma * nextpop)
            if(sum(x) <= 0){ 
              x  <- c(1, rep(0, n_s - 1))
            } else {
              x <- x / sum(x)
            }
          }
          transition[k, , i] <- x
      }
  }
  if(any(is.na(transition))) stop("error creating transition matrix")
    transition
}
```

@@

Tmax <- 500
true_distrib <- prob_dynamics(m_true[,,2], X, Tmax)
believe_distrib <- prob_dynamics(m_belief[,,2], X, Tmax)
bind_rows(true = true_distrib, believe = believe_distrib, .id = "model") %>% 
  ggplot(aes(state,probability, col=model))  + geom_line(alpha=0.5) # + geom_bar(stat="identity")
```

```{r}
x0 <- which.min(abs(states - 25))
```

```{r}
tibble(state = states,
       belief = actions[belief_p$policy],
       ideal =  actions[true_p$policy]) %>%
  pivot_longer(c(belief, ideal), 
               names_to="model",
               values_to = "action") %>%
  ggplot(aes(state,action, color = model)) + geom_point(alpha=0.5) 
```

```{r}
set.seed(12345)
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = x0, Tmax = 100)
df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

```{r}
set.seed(12345)
df <- mdp_planning(m_belief, reward, discount, model_prior = c(1), 
                   policy = belief_p$policy, x0 = x0, Tmax = 1000)

```{r}
df <- mdp_planning(m_true, reward, discount, model_prior = c(1), 
                   policy = true_p$policy, x0 = x0, Tmax = 1000)

df %>% 
  mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + 
  geom_point(alpha=0.5)  + 
  geom_line(aes(time, action), col="blue")
```

```{r}
mdp_learning_ <- memoise::memoise(mdp_learning)
  learning_sim <- mdp_learning_(transition, reward, discount, 
                      x0 = x0, 
                      Tmax = Tmax, 
                      true_transition = transition[[true_i]],
                      model_prior = prior,
                      type = "value iteration", 
                      epsilon = 1e-2)

# Simulations for Gretas

@@

@@
